Secure Composition \\ of Robust and Optimising Compilers

### 2023-09-29 csf/main.tex
Minor typo fix

\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
\end{theorem}

@@

\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. \Coqed
\end{theorem}

@@

So far, this paper assumes a unified trace-model between source and target languages.
Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs
% in such a way that the projections to the specification trace model become unequal to each other, 
% i.e., source and target language of the compiler essentially have a different trace model. 
However, there are instances of passes where this is not the case.
For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and this section re-introduces the compositionality framework with support for these situations.

To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
% This section is concerned about a compilation setting with different source and target trace models. 
These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties $\src{\class}$ to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
The section refers to these functions as existential images.

@@

For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, and \pul{relation $\sim_2$ is well-formed with respect to $\src{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, and \pul{$\wf{\sim_2}{\src{\class[_1]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

>>>>

\begin{theorem}[Upper Composition of Secure Compilers]\label{thm:urtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}+\obj{L}}{\trg{L}}}{\class[_{1}]\cap\class[_{2}]}$}. %\Coqed
\end{theorem}

@@

\begin{theorem}[Lower Composition of Secure Compilers]\label{thm:lrtp}
  If \bul{$\rtp{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}$} and \rul{$\rtp{\cc{\obj{L}}{\trg{L}}}{\class[_{2}]}$}, then \oul{$\rtp{\cc{\src{L}}{\trg{L}+\obj{L}}}{\class[_{1}]\cap\class[_{2}]}$}. % \Coqed
\end{theorem}

@@

So far, this paper assumes a unified trace-model between source and target languages.
Moreover secure compilers (in the sense of \Cref{def:rtp}) cannot change the behavior of compiled programs.
It is worth noting that, as our case study in \Cref{sec:casestud:defs} demonstrates, several compilation passes do not change the behavior of compiled programs.
% in such a way that the projections to the specification trace model become unequal to each other, 
% i.e., source and target language of the compiler essentially have a different trace model. 
However, there are instances of passes where this is not the case.
For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and this section re-introduces the compositionality framework with support for these situations.

To reason about these kinds of compilers, we rely on existing work~\cite{abate2021extacc} that lifts the presented secure compilation theory to compilers whose source and target trace models are different.
% These compilers require a Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
% This section is concerned about a compilation setting with different source and target trace models. 
These compilers require a way to say that two traces $\src{\trace}$ and $\trg{\trace}$ are related ($\sim$) and a function to translate a source-level property $\src{\pi}$ into a target-level property ($\tau_\sim\left( \src{\pi} \right)$~\cite{abate2021extacc}) using this trace relation.
The function to map source-level properties to target-level properties can be lifted to translate source-level classes of properties ($\src{\class}$) to target-level classes ($\tilde{\tau}_\sim\left(\src{\class}\right)$).
The section refers to these functions as existential images.

@@

For the adapted sequential composition theorem, let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\src{\class[_1]}$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\tilde{\tau}_{\sim_1}\left(\src{\class[_2]}\right)$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_1$ is well-formed with respect to class $\src{\class[_2]}$}, and \pul{relation $\sim_2$ is well-formed with respect to $\src{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\src{\class[_1]}\cap\src{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed with respect to $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
  If \bul{$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class[_{1}]}}{\sim_1}$}, \rul{$\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\src{\class[_{2}]}\right)}{\sim_2}$}, and \iul{$\wf{\sim_1}{\src{\class[_2]}}$}, and \pul{$\wf{\sim_2}{\src{\class[_1]}}$}, then \oul{$\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\src{\class[_{1}]}\cap\src{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}
